---
title: "Workout1"
author: "Danny Xia"
date: "9/26/2018"
output: github_document
---
#Workout 1 - Data Wrangling and Visualization

```{r}
library(dplyr)
library(ggplot2)
```

##Ranking of Teams
```{r barplot of avg salary}
teams = read.csv("../data/nba2018-teams.csv")
team_sal = arrange(teams, desc(salary))
avg_salary = mean(teams$salary)
ggplot(team_sal, aes(x = reorder(team, 30:1), y = salary)) +
  geom_bar(aes(fill = salary), stat = "identity") +
  coord_flip() +
  geom_abline(intercept = avg_salary, slope = 0, color = "red") +
  labs(x = "Team", y = "Salary (in millions)", title = "NBA Teams Ranked by Total Salary")
```

```{r barplot of total points}
team_pts = arrange(teams, desc(points))
avg_pts = mean(teams$points)
ggplot(team_pts, aes(x = reorder(team, 30:1), y = points)) +
  geom_bar(aes(fill = points), stat = "identity") +
  coord_flip() +
  geom_abline(slope = 0, intercept = avg_pts, color = "red") +
  labs(x = "Team", y = "Points", title = "NBA Teams Ranked by Total Points")
```

```{r barplot of efficiency}
team_eff = arrange(teams, desc(efficiency))
avg_eff = mean(teams$efficiency)
ggplot(team_eff, aes(x = reorder(team, 30:1), y = efficiency)) +
  geom_bar(aes(fill = efficiency), stat = "identity") +
  coord_flip() +
  geom_abline(slope = 0, intercept = avg_eff, color = "red") +
  labs(x = "Team", y = "Efficiency", title = "NBA Teams Ranked by Total Efficiency")
```

```{r barplot of Score}
team_score = 
  teams %>% 
  mutate(score = 1.5*points3 + points2 + points1 + 1.5*assists + salary - turnovers)
team_score = arrange(team_score, desc(score))
avg_score = mean(team_score$score)
ggplot(team_score, aes(x = reorder(team, 30:1), y = score)) +
  geom_bar(aes(fill = score), stat = "identity") +
  coord_flip() +
  geom_abline(slope = 0, intercept = avg_score, color = "red") +
  labs(x = "Team", y = "Score", title = "NBA Teams Ranked by Score")
```

+ The statistic that I have come up with, named Score, is
\[
1.5*points3 + points2 + points1 + 1.5*assists + salary - turnovers
\]
  + I have weighted 3 points made and assists because there aren't as many per game as the other statistics like free throws and 2-point field goals, but teams with high 3-pointers and assists in the current NBA tend to perform better. Rebounds don't always indicate a stronger team, since many good teams look for higher quality shots and do not miss as often, so there are less chances to get rebounds. These trends can be easily seen by teams like the Warriors, Rockets, and Cavaliers. Salary is also a large factor, as teams that go over the cap tend to do so because they have a strong team that they do not want to separate in spite of the salary cap and luxury tax and other NBA cap rules. Also, salary correlates with performance. Finally, limiting turnovers is an essential part of how well a team plays, given that possessions are opportunities to score, and a turnover is losing that opportunity and giving one more to an opponent.

##Comments and Reflections
1. No, this was not my first time working on a project with such a file structure.
2. No, this was not my first time using relative paths. They are important for reproducibility purposes because they absolute paths differ between people, and relative paths, given the same file structure, can be reproduced.
3. No, this was not my first time using an R script. It is definitely jarring to write code without markdown syntax, though.
4. It was hard to remember how to clean up the data with dplyr at times. It was also difficult to remember ggplot syntax, and bash commands to update the github repository. 
5. Writing the code and descriptions and basically everything else was relatively easy, and most of the things we did were done in class. Learning how to make a horizontal graph was not difficult with the link.
6. No one helped me in completing this assignment.
7. The most time consuming part was writing the data dictionary, or at least it felt like an eternity.
8. The most interesting part was creating visuals.